% -*- compile-command: "pdflatex 02-GCD-analysis-plan.tex" -*-
\documentclass{tufte-handout}

\usepackage{../algo-activity}

\title{\thecourse\ Activity 2: GCD Analysis (Facilitation Plan)}
\date{}

\begin{document}

\maketitle

Roles: get the role card you had last class, then hand it to your
right.  Take a minute to review.

\section{Learning Objectives}

\subsection{Content objectives}

\begin{itemize}
\item Students will explain the benefits of a POGIL classroom.
\end{itemize}

\subsection{Process objectives}

\section{Process Skills and Guided Inquiry (20m)}

\textbf{3 minutes} for section 1!  Report out---discussion.  5 mins total.

\emph{guided inquiry}---lots of evidence shows that you learn much
better when you have to work through things yourself rather than
listening to someone talk.

\emph{Next, they will do an exercise about qualities they look for in
  people to hire, and pick which qualities are important.  Go around
  and have each reporter add one thing that hasn't already been said,
  write each on the board.} (This isn't actually that helpful.  Maybe
have each group pick one most important thing that hasn't been said
and have them report on that?)

These are what POGIL calls \emph{process skills}.  We're going to
explicitly practice these skills by working in learning teams and
taking on different roles---that's one reason for this format.

\section{GCD (30m)}

\begin{itemize}
\item Review questions: 10 mins or so.  Discussion re: 5 mod 0.

\item After Model 1 (8 mins), just share out very quickly to make sure
  everyone is on the same page.  If time, have reflectors share one
  strength of group and one thing to improve.\marginnote{F17, didn't
    do this reflector thing, but it probably would have been a good
    idea\dots}

\item Model 2 (10 mins).

\item Closure (2 mins): hopefully you have started to develop
  intuitions as to why these algorithms work.  Next time we will turn
  these intuitions into formal proofs.  Note HW 1 posted on website.
  Get started early!  We will review proof by induction and see
  examples on Monday.

\end{itemize}


\end{document}
