% -*- compile-command: "pdflatex --enable-write18 09-divide-and-conquer-practice-plan.tex" -*-
\documentclass{tufte-handout}

\usepackage{../algo-activity}
\usepackage{../acode}

\title{\thecourse\ Activity 9: Divide \& Conquer Practice
  (Facilitation Plan)}
\date{}

\begin{document}

\maketitle

Remember the 3 components of divide \& conquer:
\begin{itemize}
\item \emph{Divide} input into subproblems somehow (split, partition, \dots)
\item \emph{Recursively} solve the subproblems (\emph{have faith} they
  are correct!)
\item \emph{Combine} the answers
\end{itemize}

Don't forget base case(s) as well.

To prove (by induction on size of input!):
\begin{itemize}
\item Prove base case(s) are correct
\item Prove: \emph{if} answers to subproblems are correct, \emph{then}
  answer to whole problem is correct.
\end{itemize}

\section{Array Median}

\noindent
\textbf{Input}: array of $n$ integers. \\
\noindent
\textbf{Output}: the median element, \emph{i.e.} the $\lfloor n/2
\rfloor$-smallest. \bigskip

\textbf{In groups}: what's the first algorithm you think of?  How long
does it take? (Probably sort + return middle, $\Theta(n \lg n)$.)

Can we do better?  Get ideas.

Problem: if we \emph{e.g.} split the array in half somehow and get the
median of both, that wouldn't really help.  The medians of the
subarrays are not likely to have any relationship to the median of the
whole.

Solution: generalize!  Let's find not just the median, but take $k$ as
an input and return the $k$th-smallest.  $\textsc{Select}(A,k)$:
return the $k$th-smallest element of array $A$.  We can get median by
calling $\textsc{Select}(A, \lfloor n/2 \rfloor)$.

Question: how should we split up the array?  Just splitting down the
middle doesn't really help.  What's a different way we could split up
$A$ (think in groups). Hint: think of other divide \& conquer
algorithms you know. (Quicksort.)

Solution: pick a pivot element $p$ (random!) and partition into $A_1$
= all elements $<p$
and $A_2 =$ all elements $\geq p$.  Then if $k < |A_1|$ return
$\textsc{Select}(A_1, k)$, else return $\textsc{Select}(A_2, k - |A_1|)$.

Prove correct.  Analyze time complexity.

If time left, think about maximal subarray problem.

\end{document}
