% -*- compile-command: "pdflatex AA-plan.tex" -*-
\documentclass{tufte-handout}

\usepackage{algo-activity}

\title{\thecourse: Asymptotic Analysis (Facilitation Plan)}
\date{}

\begin{document}

\maketitle

\section{Announcements}

\begin{itemize}
\item Remember HW 1 due Friday.  Start early, come ask for help if you
  need it.
\item Assign groups?

\item Today, take a role you haven't had.  Review duties.
\end{itemize}

\section{Learning Objectives}

\subsection{Content objectives}

\begin{itemize}
\item Students will describe asymptotic behavior of functions using
  big-$O$, big-$\Theta$, and big-$\Omega$ notation.
\end{itemize}

\subsection{Process objectives}

\section{CTQs I (Big-O) (32 mins)}

(Up to 3 minutes to get started, look at role cards, etc.)

\begin{itemize}
\item Make grid, go around and tell them put up answers to T/F
  questions once they get there.  Discuss as necessary.
\item Also tell reflectors I will be asking them to share insights
  with group later.  One strength, one improvement.
\end{itemize}

\section{Reflection (1:42, 2 mins)}

Have reflectors take 2 mins to share one strength and one thing to
improve with their group.

\begin{reflect}{F'18}
  Actually did it this time, and I think it was really great.  Going
  to try to use reflectors more intentionally this semester.
\end{reflect}

\section{Model 2 (definitions, limits) (15 mins)}

Just let them work, observe.  Hopefully they will get to the end.

\section{Closure (1 min)}

If you didn't get to the end, finish on your own, discuss at
beginning of next class.

\begin{reflect}{F'18}
  Due to time constraints, had them skip writing their own
  definitions, and focus on questions 16 and 17 (which have them
  connect the formal definitions back to their intuition).  Didn't
  really get to the rest of the questions (about doing formal proofs
  using definitions) but that's totally OK.  Overall I think it went
  super well, I could tell that lots of learning was happening.
\end{reflect}

\end{document}
