% -*- compile-command: "pdflatex role-cards.tex" -*-

\documentclass[12pt]{article}

\pagestyle{empty}

\usepackage{xspace}
\usepackage[landscape]{geometry}

\newcommand{\blank}{\ensuremath{\underline{\phantom{XXXXXX}}}\xspace}

\begin{document}

\section*{Manager}

You are the \textbf{manager} for today.  Your job is to ensure that
team members are fulfilling their roles, the assigned tasks are being
accomplished on time, and all members of the team are participating
and contributing. Specific responsibilities include:

\begin{itemize}
\item Make sure the team starts quickly and remains focused during the
  activity.
  \begin{itemize}
  \item ``Are we ready to begin?''
  \item ``I think we're on a bit of a tangent, let's get back to
    working on this.''
  \end{itemize}
\item Take care of time management.
  \begin{itemize}
  \item Keep an eye on the clock.
  \item Keep the team moving forward and communicate about discussion
    deadlines.
  \item ``I think we need to focus on \blank so we complete this
    section on time.''
  \item ``We have \blank minutes before we need to discuss this. Let's
    get this done.''
  \end{itemize}
\item Make sure	all voices on the team are heard.
  \begin{itemize}
  \item Address team members by name and ensure that everyone
    contributes.
  \item Assign different members to read sections of the activity on a
    rotating basis.
  \item ``(Name), what do you think about \dots?''
  \item ``I would like to hear what you think, (name).''
  \end{itemize}
\item Help make sure other team members are fulfilling their roles.
  \begin{itemize}
  \item ``(Reflector name), do you have any thoughts about how our
    team is doing so far?''
  \item ``(Recorder name), could you write that down?''
  \item ``(Reporter name), do you know what you're going to say, or do
    you need more input from the rest of us?''
  \end{itemize}
\end{itemize}

\newpage
\section*{Recorder}

You are the \textbf{recorder} for today. Your job is to ensure that
the team's work is recorded accurately for later reference. Although
all team members may be writing down some answers and conclusions on
their own copy of the activity, you should make sure that yours is
particularly complete.  Specific responsibilities include:

\begin{itemize}
\item Record the names and roles of the group members at the
  beginning of each activity.
\item Record the important aspects of group discussions, observations,
  insights, etc.
  \begin{itemize}
  \item ``This seems like an important conclusion. Let's stop
    for a minute so I can write it down.''
  \item ``That was a great insight, (name). Do you mind if I quote you
    in our report?''
  \item ``Can you repeat that more slowly so I can write it down?''
  \end{itemize}
\item Completely fill out your group's answers for the day's activity.
\item After class, you should scan your notes and distribute copies to
  your team members.
\end{itemize}

\newpage
\section*{Reporter}

You are the \textbf{reporter} for today.  Your job is to handle
communication between your team and other teams or the instructor.
Specific responsibilities include:

\begin{itemize}
\item Communicate team questions and clarifications with the instructor
  or other teams.  (You are the \textbf{only} team member who should
  do so.)
  \begin{itemize}
  \item ``Our team is confused about how \blank relates to \blank.''
  \end{itemize}
\item Ensure all team members have had a chance to respond before
  asking outside sources.
  \begin{itemize}
  \item ``Does anyone in our team know the answer to \blank?''
  \item ``Before we ask the teacher, could someone clarify the answer
    to \dots''
  \end{itemize}
\item Ensure that everyone in the team agrees on what to ask if an
  outside source is needed.
  \begin{itemize}
  \item ``Does everyone agree we need to find out \dots?''
  \end{itemize}
\item Present conclusions of the team to other groups or the class, as
  requested.
  \begin{itemize}
  \item ``Our team reached consensus that the answer to number \blank was \blank.''
  \item ``The reasoning we used to answer number \blank was \dots''
  \end{itemize}
\end{itemize}

\newpage
\section*{Reflector}

You are the \textbf{reflector} for today.  Your job is to observe team
dynamics with respect to the learning process, to point out strengths
and potential improvements, and to help guide the consensus-building
process.  You will also give feedback to the instructor on both your
team and the activity.

Specific responsibilities include:

\begin{itemize}
\item Observe team dynamics and behavior with respect to the learning process.
  \begin{itemize}
  \item Is everyone in the team participating?
  \item Are team members listening carefully to each other?
  \item Are team members being patient and respectful of each other?
  \end{itemize}
\item Guide the consensus-building process; your team must agree on
  responses to questions.
  \begin{itemize}
  \item ``Would you all agree that \blank is a good answer for
    question number \blank?''
  \item ``Could you please rephrase what you just said?''
  \item ``Is your answer completely supported by your explanation?''
  \item ``Would that response make sense to someone from another
    team?''
  \end{itemize}
\item Report to the team periodically during the activity on how the
  team performs.
  \begin{itemize}
  \item ``Let's stop for a minute. I have a couple comments on what we
    are doing well and a suggestion on how we could be more
    productive.''
  \end{itemize}
\item Give feedback to the instructor (via a short online survey)
  regarding the activities and your team's learning process.
  \begin{itemize}
  \item You are welcome to get input from other team members in
    putting together your feedback.
  \item ``What do you think was a strength of this activity?''
  \item ``Does anyone have any suggestions for how this activity could
    be improved?''
  \end{itemize}
\end{itemize}

\end{document}
